### 2014-09-12 tablewithgraphs.tex
add table with graphs

\begin{longtable}{ c || c | c | c | c }
\hline
matrix & connectivity & cycle number & \specialcell{probability of stability\\to perturbation} & probability of stability\\
\hline
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 3 & 0 & 0.499 & 0.126\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 4 & 1 & 0.505 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 4 & 0 & 0.622 & 0.127\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.595 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.494 & 0.128\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.41 & 0.061\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.43 & 0.078\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.605 & 0.12\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 2 & 0.405 & 0.06\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 0 & 0.698 & 0.122\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 1 & 0.587 & 0.128\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 5 & 0 & 0.707 & 0.127\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.578 & 0.121\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 3 & 0.487 & 0.081\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.543 & 0.098\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.501 & 0.088\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.662 & 0.123\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 3 & 0.467 & 0.079\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 2 & 0.583 & 0.13\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.659 & 0.124\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 0 & 0.751 & 0.124\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 6 & 1 & 0.604 & 0.097\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 3 & 0.564 & 0.103\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 5 & 0.475 & 0.068\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 3 & 0.591 & 0.108\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 1 & 0.717 & 0.119\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 2 & 0.648 & 0.122\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 7 & 2 & 0.627 & 0.105\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 8 & 5 & 0.577 & 0.093\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 8 & 3 & 0.639 & 0.109\\
\begin{tikzpicture}[scale=0.6,
every state/.style={draw=red!50,very thick,fill=red!20}]
\begin{dot2tex}[styleonly,codeonly,neato,mathmode]
digraph G {
d2ttikzedgelabels = true;
node [style="state"];
edge [lblstyle="auto",topath="bend left",style="line width=1.5pt"];
1 -> 1 [topath="loop above"];;
2 -> 1;
3 -> 1;
1 -> 2;
2 -> 2 [topath="loop above"];;
3 -> 2;
1 -> 3;
2 -> 3;
3 -> 3 [topath="loop above"];;
}
\end{dot2tex}
\end{tikzpicture} & 9 & 5 & 0.638 & 0.106\\
\caption{{\bf Probability of stability under resampling and \emph{a priori} stability for three component systems estimated via Monte Carlo sampling.} All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat3graph}
\end{longtable}