### 2014-09-26 tex/figuressupp.tex
add figure showing symmetries of robustness

%!TEX root = ../paper.tex
% \begin{figure}[!ht]
% \centering
% \noindent\includegraphics[width=0.7\columnwidth]{fig/region2and3.pdf}
% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two and three component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial.}
% \label{fig:region2and3}
% \end{figure}

% \pagebreak

% \begin{figure}[!ht]
% \centering
% \noindent\includegraphics[width=0.5\columnwidth]{fig/region2x2.pdf}
% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for two component systems.} The regions correspond to all possible relationships between the invariants determined by the characteristic polynomial. Colors correspond to the regions given in the main text Green: $R_{2000}$, Red: $R_{1100}$, Yellow: $R_{0200}$, Magenta: $R_{0002}$, Blue: $R_{0020}$.}
% \label{fig:region2x2}
% \end{figure}

% \pagebreak

% \begin{figure}[!ht]
% \centering
% \noindent\includegraphics[width=0.5\columnwidth]{fig/region3x3.pdf}
% \caption{{\bf Stability conditions on coefficients of the characteristic polynomial for three component systems.} Colors correspond to the following regions Green: $R_{3000}$, Red: $R_{1200}$, Yellow: $R_{1002}$, Blue: $R_{1020}$. This is the plane $s_3=1$.}
% \label{fig:region3x3}
% \end{figure}