### 2014-10-15 paper.tex
revise

\input{header.tex}

\begin{document}

% Add Figure, Table prefixes to references
% http://tex.stackexchange.com/a/6063/6784
\let\ref\autoref

\pagenumbering{arabic}

% \twocolumn[
% \begin{@twocolumnfalse}
\title{Ensemble Stability of Biological Networks}

% Title must be 150 characters or less
\begin{center}
{\Large
\textbf{Hierarchical Network Structure Promotes Dynamical Robustness}
}
% Insert Author names, affiliations and corresponding author email.
\\[.5cm]
Cameron Smith$^{1}$,
Raymond S. Puzio$^{1}$,
Aviv Bergman$^{1,2,3,4}$,
\\[.5cm]
$^1$Department of Systems and Computational Biology,\\
$^2$Dominick P. Purpura Department of Neuroscience,\\
$^3$Department of Pathology, Albert Einstein College of Medicine,\\
1301 Morris Park Ave, Bronx, NY 10461, USA\\
$^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
\\[.5cm]
% $\ast$To whom correspondence should be addressed; E-mail: aviv@einstein.yu.edu.
\end{center}
% \end{@twocolumnfalse}
% ]

% \begin{multicols}{2}

% {\begin{quote} \bf
% \input{tex/abstract-cited}
% \end{quote}}
{\bf
\input{tex/abstract-cited}
}

% \section{Todo}
% \listoftodos
% \tableofcontents

% \section{Introduction}
\input{tex/introduction}

% \section{Results}
\input{tex/results}

% \section{Discussion}
\input{tex/discussion}

\section{Acknowlegements}
\input{tex/acknowledgements}

% \pagebreak

{\footnotesize
\bibliographystyle{pnas}
\bibliography{bib/books,bib/papers}
}

% \end{multicols}

\pagebreak
\FloatBarrier
\section{Figure Legends}
\input{tex/figures}

\pagebreak
\FloatBarrier

\beginsupplement
\setcounter{secnumdepth}{4}
\section{Supplementary Material}

\input{tex/supplement}
\pagebreak

\input{tex/figuressupp}
\pagebreak

% \section{Tables}
% \todo{Fill out notation table}
% \input{tex/tablenotation}
\input{tex/table2x2stab}
\pagebreak
\input{tex/tablenographs}

% \input{tex/tablewithgraphs}
% \input{tex/extra}

\end{document}

%------------------------------------------------------------------------------
% End of journal.tex
%------------------------------------------------------------------------------