### 2014-11-26 paper.tex
update text

\input{header.tex}

\begin{document}

% Add Figure, Table prefixes to references
% http://tex.stackexchange.com/a/6063/6784
\let\ref\autoref

\pagenumbering{arabic}

\title{\bf Hierarchical Network Structure Promotes Dynamical Robustness}

\author{Cameron Smith$^{1}$,
Raymond S. Puzio$^{1}$,
Aviv Bergman$^{1,2,3,4,*}$}

\affiliation{$^1$Department of Systems and Computational Biology,\\
$^2$Dominick P. Purpura Department of Neuroscience,\\
$^3$Department of Pathology, Albert Einstein College of Medicine,\\
1301 Morris Park Ave, Bronx, NY 10461, USA\\
$^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
\\
$*$to whom correspondence should be addressed: aviv@einstein.yu.edu}

\date{\today}

\input{tex/abstract-cited}

\maketitle

% \section{Todo}
% \listoftodos
% \tableofcontents

\section{Introduction}
\input{tex/introduction}

\section{Results}
\input{tex/results}

\section{Discussion}
\input{tex/discussion}

\section{Methods}
{\footnotesize
\input{tex/methods}
}

\section{Acknowlegements}
\input{tex/acknowledgements}

\bibliographystyle{unsrtnat}
\bibliography{bib/books,bib/papers}

\onecolumngrid

\pagebreak
\FloatBarrier

\beginsupplement
\setcounter{secnumdepth}{4}
\section{Supporting Information}

\input{tex/figuressupp}
\pagebreak

% \section{Tables}
\input{tex/table2x2stab}
\pagebreak
\input{tex/tablenographs}

\end{document}

%------------------------------------------------------------------------------
% End of paper.tex
%------------------------------------------------------------------------------