### 2015-06-03 paper.tex
turn on section numbering

\input{header.tex}

\begin{document}

% Add Figure, Table prefixes to references
% http://tex.stackexchange.com/a/6063/6784
\let\ref\autoref

% Remove prefix of footnote
% \makeatletter
% \renewcommand{\thempfn}{\thefootnote}
% \makeatother

\pagenumbering{arabic}

\title{\bf Hierarchical Network Structure Promotes Dynamical Robustness}

\author{Cameron Smith$^{1}$,
Raymond S. Puzio$^{1}$,
Aviv Bergman$^{1,2,3,4,*}$}

\affiliation{$^1$Department of Systems and Computational Biology,\\
$^2$Dominick P. Purpura Department of Neuroscience,\\
$^3$Department of Pathology, Albert Einstein College of Medicine,\\
1301 Morris Park Ave, Bronx, NY 10461, USA\\
$^4$Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA
\\
$*$to whom correspondence should be addressed: aviv@einstein.yu.edu}

\date{\today}

\input{tex/abstract}

\maketitle

% \section{Todo}
% \listoftodos
% \tableofcontents

\begin{bibunit}[pnas]

% comment to omit numbering of sections
\setcounter{secnumdepth}{4}

% \section{Introduction}
\input{tex/introduction}

% \section{Results}
% \input{tex/results}
\input{tex/overview}

% \section{Discussion}
\input{tex/discussion}

% \section{Methods}
% {\footnotesize
% \input{tex/methods}
% }

\section*{Acknowlegements}
\input{tex/acknowledgements}

% \bibliographystyle{Science}
% \bibliography{bib/books,bib/papers}

\putbib[bib/books,bib/papers]

\pagebreak
\FloatBarrier

\onecolumngrid
\pagebreak
% \section{Figures}
\input{tex/figures}

\end{bibunit}
\pagebreak
\twocolumngrid

\pagebreak
\FloatBarrier

\beginsupplement
\setcounter{secnumdepth}{4}

\begin{bibunit}[unsrtnat]

\begin{widetext}
\begin{center}
\text{\large \refsupp{} for }\\
\textbf{\large Hierarchical Network Structure Promotes Dynamical Robustness}\\
\text{Cameron Smith, Raymond S. Puzio, Aviv Bergman$^*$}\\
\text{$^*$Corresponding author. E-mail: aviv@einstein.yu.edu}
\end{center}
\end{widetext}

% \section{\refsupp{}}

\input{tex/results}
\input{tex/methods}

\putbib[bib/books,bib/papers]

\end{bibunit}

\pagebreak
\onecolumngrid
\pagebreak

\input{tex/figuressupp}
\pagebreak
\FloatBarrier

% \section{Tables}
\input{tex/table2x2stab}
\pagebreak
\input{tex/tablenographs}

\end{document}

%------------------------------------------------------------------------------
% End of paper.tex
%------------------------------------------------------------------------------