### 2015-07-27 tex/table2x2stab.tex
update explanation of relative contribution of SCCs to robustness

%!TEX root = ../paper.tex
% \onecolumn
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\begin{table*}[h]
\begin{center}
\begin{tabular}{ c || c | c | c | c | c }
\hline
\specialcell{adjacency \\ matrix} & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{$\frac{2d_i}{v_i}(1-R_i)$ \\ (\ref{eq:maxrobcond})}\\
\hline
  $\begin{pmatrix}
1 & 1 \\
1 & 1
\end{pmatrix}$ & 1 & 4 & 0.62 & 0.25 & 1.52\\
  $\begin{pmatrix}
1 & 1 \\
1 & 0
\end{pmatrix}$, $\begin{pmatrix}
0 & 1 \\
1 & 1
\end{pmatrix}$ & 1 & 3 & 0.5 & 0.25 & 1.5\\
  $\begin{pmatrix}
1 & 0 \\
1 & 1
\end{pmatrix}$, $\begin{pmatrix}
1 & 1 \\
0 & 1
\end{pmatrix}$ & 2 & 3 & 0.67 & 0.25 & \\
$\begin{pmatrix}
1 & 0 \\
0 & 1
\end{pmatrix}$ & 2 & 2 & 0.5 & 0.25 & \\
\end{tabular}
\end{center}
\caption{{\bf Dynamical robustness and probability of stability for two variable systems derived analytically}. All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
\end{table*}
% \twocolumn